---
title: "Anexo Homework 4"
author: "Gerardo Mamani"
date: "10/13/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r, comment=NA}
pedigree<-read.table(header=T, text = 'ID sire dam gener
                                Lord_Raglan 0 0 1
                                Mistletoe Lord_Raglan 0 2
                                Champion_of_England 0 0 3
                                Duchess_of_Gloster Lord_Raglan 0 2
                                The_Czar Lord_Raglan 0 2
                                Mimulus Champion_of_England Mistletoe 3
                                Grand_Duke_Gloster Champion_of_England Duchess_of_Gloster 3
                                Carmine The_Czar 0 3
                                Royal_Duke_Gloster Grand_Duke_Gloster Mimulus 4
                                Princess_Royal Champion_of_England Carmine 4
                                Roan_Gauntlet Royal_Duke_Gloster Princess_Royal 5 ')
                                
pedigree
```

### Recoding the pedigree names by numbers

Id   | code     
---- | ---- 
Lord Raglan    | 1   |       
Mistletoe     | 2            
Champion of England    | 3       
Duchess of Gloster    | 4       
The Czar    | 5    
Mimulus    | 6      
Grand Duke Gloster    | 7      
Carmine    | 8    
Royal Duke Gloster    | 9      
Princess Royal    | 10     
Roan Gauntlet    | 11      

### Working with pedigree recoded

```{r}
pedigree<-read.table(header=T, text = 'Id sire dam gen
                                       1	0	0	1
                                       2	1	0	2
                                       3	0	0	2
                                       4	1	0	2
                                       5	1	0	2
                                       6	3	2	3
                                       7	3	4	3
                                       8	5	0	3
                                       9	7	6	4
                                       10	3	8	4
                                       11	9	10 5 ')
               
```

### Using synbreed package

```{r}
library(synbreed)
pedigree <- create.pedigree(pedigree$Id,pedigree$sire,pedigree$dam,pedigree$gen,unknown=c('Id','0'))
```

```{r, warning=NA, comment=NA}
plot(pedigree)
```
\newpage
```{r, comment=NA}
A <- kin(pedigree,ret="add")
round(A,5)
```


\newpage

## Question 3

### First case, when parents of individual B are full-siblings relative

```{r}
ped3a<-read.table(header=T, text = 'Id sire dam ger
                                     M	0	0	1
                                     N  0 0 1  
                                     X	M	N	2
                                     Y	M	N	2
                                     A	0	0	3
                                     B	X	Y	3
                                     C	A	B	4
                                     D	A	B	4
                                     E	C	D	5 ')
               
```

```{r, comment=NA}
library(synbreed)
ped3a <- create.pedigree(ped3a$Id,ped3a$sire,ped3a$dam,ped3a$ger,unknown=c('Id','0'))
ped3a
```

```{r, comment=NA, warning=TRUE}
plot(ped3a)
```

\newpage
```{r, comment=NA}
A3a <- kin(ped3a,ret="add")
round(A3a,5)
```

### Second case, when the parents of B are unrelated

```{r cars}
ped3b<-read.table(header=T, text = 'Id sire dam ger
                                     X	0	0	1
                                     Y	0	0	1
                                     B	X	Y	2
                                     A	0	0	2
                                     C	A	B	3
                                     D	A	B	3
                                     E	C	D	4')
```

```{r, comment=NA}
library(synbreed)
ped3b <- create.pedigree(ped3b$Id,ped3b$sire,ped3b$dam,ped3b$ger,unknown=c('Id','0'))
ped3b
```

\newpage
```{r, comment=NA, message=FALSE}
plot(ped3b)
```

```{r, comment=NA}
A3b <- kin(ped3b,ret="add")
round(A3b,5)
```


