---
title: "Agro/ANSCI 931 Population Genetics"
author: "Gerardo Mamani"
date: "Lincoln, 11/1/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 1. A bi-allelic locus affects oil content (\%) in rye with the following genotypic values: 

### A1A1: 9.0\% A1A2: 8.9\% A2A2: 8.0\%. 

### The frequency of A1 is 0.30 and the population is in Hardy-Weinberg Equilibrium.

#### a. Given these genotypic values, what are the values of a and d? 

mid-point = \((A_1A_1 + A_2A_2)/2 = 8.5\%\)  

a = \(A_1A_1 - mid-point = 9.0\% - 8.5\% = 0.5\)  

d = \(A_1A_2 - mid-point = 8.9\% - 8.5\% = 0.4\)  

#### b. What is the population mean?

\(M=a(p-q)+2dpq\)


```{r}
p<-0.3 # Frequency of A1
q<-0.7 # Frequency of A2
a<-0.5 # value of a
d<-0.4 # value of d
```
```{r, comment=NA}
mean<-a*(p-q)+(2*d*p*q)
mean
```

##### c. What is the mode of inheritance?

Partial dominance

### 2. Consider two more loci that affect oil content, B and C. Assume B1 and C1 have exactly the same average effect as A1. Likewise, B2 and C2 have exactly the same average effect as A2.


#### a. What is the breeding value of an individual that is \(A_1A_2B_1B_2C_1C_2\) ? 

```{r}
p<-0.3 # Frequency of A1
q<-0.7 # Frequency of A2
a<-0.5 # value of a
d<-0.4 # value of d
```

 Average value of allele \(A_1, \space  \alpha_1=q[a+d(q-p)]\)
```{r, comment=NA}
ave_A1<-q*(a+d*(q-p))
ave_A1
```

 Average value of allele \(A_2, \space  \alpha_2=-p[a+d(q-p)]\):
```{r, comment=NA}
ave_A2<--p*(a+d*(q-p))
ave_A2
```

 Breeding Value of genotype \(A_1A_2=\alpha_1 + \alpha_2\):
```{r, comment=NA}
BV_A1A2<-ave_A1+ave_A2
BV_A1A2
```

Based in the assuntion, then the breeding value of \(A_1A_2B_1B_2C_1C_2\) is:

```{r, comment=NA}
BV_A1A2B1B2C1C2=3*ave_A1 + 3*ave_A2
BV_A1A2B1B2C1C2
```


#### b. What is the breeding value of an individual that is \(A_1A_1B_1B_1C_1C_1\)?

Based in the assuntion \(A_1=B_1=C_1\), then the breeding value of \(A_1A_1B_1B_1C_1C_1 \) is:

```{r, comment=NA}
BV_A1A1B1B1C1C1=6*ave_A1
BV_A1A1B1B1C1C1
```

#### c. What does the calculated value in part b mean in terms of offspring performance of the individual in question?

For part b is \[\frac{2.772}{2}=1.386\]. The offspring should have 1.386 percentage points more oil than the average of the population.

#### d. What is the expected breeding value of an offspring derived from a mating between the two individuals in parts a and b?

```{r, comment=NA}
BV_offspring<-0.5*(BV_A1A2B1B2C1C2 + BV_A1A1B1B1C1C1)
BV_offspring
```
Offspring between \textbf{a} and \textbf{b} will be \[\frac{0.792}{2}+\frac{2.772}{2}=1.782\]

#### 3. Why is the concept of the average effect of an allele important? Why not just deal with the average effect of a genotype, rather than the average effect of an allele?

Because parent pass on their alleles and not their genotypes to the next generation, genotypes being created afresh in each generation.


### 4. Consider a large population of beef cattle that has been genotyped for Myostatin and phenotyped for two traits, fat and loin muscle area, with the following results.

Table

|Genotype   | Frequency | Fat (cm) |Loin Muscle Area (\(cm^2)\) |
|----       | ----      | ----     | ----                    |
|AA         | 0.5       | 0.51     | 116.1                   |
|Aa         | 0.4       | 0.76     | 103.0                   |
|aa         | 0.1       |1.5       | 83.9                    |




#### a. Compute the expected value of Fat.

|Genotype   | Frequency | Fat (cm) | Loin Muscle Area (\(cm^2)\) |Freq $\times$ Value Fat | Freq $\times$ Value Loin 
|------        | ------       | -------    | ------                        | ------            | ------ |
|AA         | 0.5       | 0.51     | 116.1     | 0.255  |58.05 |
|Aa         | 0.4       | 0.76     | 103.0     | 0.304  |41.2 |          
|aa         | 0.1       |1.5       |83.9       | 0.15  |8.39 |
|           |           |          |      Sum     | 0.709|107.64|

\(E(Fat)=\sum \space Frequency \times Genotypic \space value\)

\(E(Fat)=0.5\times0.51 + 0.4\times0.76 + 0.1\times1.5=0.709\)

#### b. Compute the variance of trait Fat

\(Var(Fat)=E(Fat^2)-E(Fat)^2\)

or it can be estimated directly as below

\(Var(Fat)=0.5(0.51-0.709)^2+0.4(0.76-0.709)^2+0.1(1.5-0.709)^2\)

\(Var(Fat)=0.083409 \space cm^2\)


#### c. Compute the expected value of Fat given that the individual carries at least one ‘a’ allele?

\(Pr(Aa, a)/Pr(a)=\frac{0.4}{0.5}=0.8\)

\(Pr(aa, a)/Pr(a)=\frac{0.1}{0.5}=0.2\)

\(E(Fat/a)=0.8 \times 0.76 + 0.2\times1.5=0.908 \space cm\)

#### d. Compute the correlation between Fat and Loin Muscle Area?

```{r, comment=NA}
fat<-c(0.51, 0.76, 1.5)
mus<-c(116.1, 103, 83.9)
cov<-(sum((fat-mean(fat))*(mus-mean(mus))))/(length(fat)-1)
var1<-(sum((fat-mean(fat))^2))/(length(fat)-1)
var2<-(sum((mus-mean(mus))^2))/(length(mus)-1)
r<-cov/sqrt(var1*var2)
r
```

\(r_{fat.mus}=\frac{Cov(Fat, Mus)}{\sqrt{Var(Fat)Var(Mus)}}=\frac{-8.2145}{\sqrt{0.2650333 \times 262.21}}=-0.9853857\)

#### e. Compute the coefficient of regression of Fat on Loin Muscle Area?

```{r, comment=NA}
cov<-(sum((fat-mean(fat))*(mus-mean(mus))))/(length(fat)-1)
var2<-(sum((mus-mean(mus))^2))/(length(mus)-1)
b<-cov/var2
b
```

\(b_{fat.mus}=\frac{Cov(Fat, Mus)}{\sqrt{Var(Mus)}}=\frac{-8.2145}{\sqrt{262.21}}=-0.03132794\)

#### f. Compute the predicted value for Loin Muscle Area for an individual from this population that has Fat=1.0?

```{r, comment=NA}
data<-data.frame(fat,mus)
mean(data$fat)
mean(data$mus)
lm(mus~fat)
```

\(E(Mus|Fat=1)= a + b.fat\)

\(E(Mus|Fat=1)=  129.6 -30.99 \times 1 = 98.63 \space cm^2\)

#### g. Compute the variance of a new variable Z, which is equal to Z = 3*Fat – 0.5*LoinMuscleArea

\(Z=3*Fat - 0.5*Mus\)

```{r, comment=NA}
Z<-3*fat - 0.5*mus
Z
```

Values for Z by genotype are :

\(AA= -56.52\)

\(Aa= -49.22\)

\(aa= -37.45\)

```{r, comment=NA}
var_Z<-0.5*(-56.52- mean(Z))^2+0.4*(-49.22- mean(Z))^2+0.1*(-37.45- mean(Z))^2
var_Z<-0.5*(-56.52- (mean(Z)))^2+0.4*(-49.22- (mean(Z)))^2+0.1*(-37.45- (mean(Z)))^2
var_Z
```

\(Var(Z)=0.5(0.51-(-47.73))^2+0.4(0.76-(-47.73)^2+0.1(1.5-(-47.73)^2\)

\(Var(Z)=50.08793 \)

#### h. Compute Cov(Z, Fat)

```{r, comment=NA}
cov_Zfat<-(sum((Z-mean(Z))*(fat-mean(fat))))/(length(Z)-1)
cov_Zfat
```

### 5. Assume that passing a course, denoted as P, is a random process that follows a Bernoulli distribution and that the \(Pr(P)=0.7\).

#### a. What is \(E(P)\) (1pt.)

\(E(P)=0.7\)

#### b. What is \(var(P)\) (1pt.)

\(0.7*0.3=0.21\)

